EDA

### 2024-11-24 other/EDA.qmd
added intro table

kable(head(df, 5), caption = "Sample of the Data")
```

@@

df <- df %>%
  mutate(year_month = floor_date(award_date, "month"))

# Aggregate the amount by 'year_month'
monthly_spending <- df %>%
  group_by(year_month) %>%
  summarise(total_amount = sum(amount, na.rm = TRUE))

@@

@@

df$amount <- as.numeric(df$amount)

@@

```{r fig-template1}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center

df <- df %>%
  mutate(contract_length_days = as.numeric(difftime(end_date, start_date, units = "days")))

# Group by buyer and calculate average length of contract and average contract size
buyer_stats <- df %>%
  group_by(buyer) %>%
  summarise(
    `Average Length of Contract (Days)` = mean(contract_length_days, na.rm = TRUE),
    `Average Contract Size` = mean(amount, na.rm = TRUE)
  )

# Reshape the data to have buyers as columns and metrics as rows
buyer_stats_long <- buyer_stats %>%
  pivot_longer(cols = -buyer, 
               names_to = "Metric", 
               values_to = "Value")

# Now pivot wider to have buyers as columns
buyer_stats_wide <- buyer_stats_long %>%
  pivot_wider(names_from = buyer, values_from = Value)

# Optionally, format the numbers
buyer_stats_wide <- buyer_stats_wide %>%
  mutate(across(-Metric, ~ round(.x, 2)))

# Display the table using knitr::kable
kable(buyer_stats_wide, caption = "Average Length of Contract and Average Contract Size by Buyer")

@@

>>>>

df_table <- df %>% select(-contract, -start_date, -end_date)

kable(head(df_table, 5), caption = "Sample of the Data") %>%
  kable_styling(font_size = 10, latex_options = c("scale_down"))

@@

df_fig <- df %>%
  mutate(year_month = floor_date(award_date, "month"))

# Aggregate the amount by 'year_month'
monthly_spending <- df_fig %>%
  group_by(year_month) %>%
  summarise(total_amount = sum(amount, na.rm = TRUE))

@@

```
```{r}

@@

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: "Average Contract Length (Days) by Government Department (Top 10)"
#| fig-align: center

# Load necessary libraries
library(dplyr)
library(ggplot2)

# Calculate average contract length per buyer (government department)
avg_contract_length <- df %>%
  mutate(contract_length_days = as.numeric(difftime(end_date, start_date, units = "days"))) %>%
  group_by(buyer) %>%
  summarise(`Average Length of Contract (Days)` = mean(contract_length_days, na.rm = TRUE)) %>%
  arrange(desc(`Average Length of Contract (Days)`)) %>%
  slice_head(n = 10)  # Select top 10 departments

# Create the bar chart
ggplot(avg_contract_length, aes(x = reorder(buyer, `Average Length of Contract (Days)`), y = `Average Length of Contract (Days)`)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(x = "Government Department", y = "Average Contract Length (Days)", title = "Average Contract Length (Days) by Government Department (Top 10)") +
  theme_minimal()

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: "Average Contract Size by Government Department (Top 10)"
#| fig-align: center

# Load necessary libraries
library(dplyr)
library(ggplot2)
library(scales)  # For formatting currency

# Calculate average contract size per buyer (government department)
avg_contract_size <- df %>%
  group_by(buyer) %>%
  summarise(`Average Contract Size` = mean(amount, na.rm = TRUE)) %>%
  arrange(desc(`Average Contract Size`)) %>%
  slice_head(n = 10)  # Select top 10 departments

# Create the bar chart
ggplot(avg_contract_size, aes(x = reorder(buyer, `Average Contract Size`), y = `Average Contract Size`)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  coord_flip() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  labs(x = "Government Department", y = "Average Contract Size", title = "Average Contract Size by Government Department (Top 10)") +
  theme_minimal()
